---
title: "Example for paper"
author: "Jan Schneider"
date: "7 8 2021"
output: pdf_document
---

## Example of asp21bpsplines

```{r results = "hide", message = FALSE}
require(asp21bpspline)
set.seed(2)
n = 1000
x = c(runif(n, 0, 20))
mean = -0.04*x^2 + 0.5* x 
y =  mean + rnorm(n, 0, 0.2 *x + 0.3)
m1 = list(x = x, z = x, y = y)

```
Consider you have the above data and want to fit a spline to it. With the 
asp21bpsplines package you would first call the spline function:

```{r}
m = spline(m1, kn = c(40, 40), order = c(3,3), p_order = c(3, 3), 
           smooth = c(1, 1))
summary(m)
plot(m)
```

# add plot as separate file

The model estimates separate parameters for the location and the scale of the data.
In this plot the middle line represents the prediction for the mean value
and the outer lines represent the 95 % confidence interval for new predictions.
However it seems that the predictions are overfitting a bit. One could now raise 
the values of the smoothing parameters in the call to the spline function. 
An alternative is a call to the MCMC function.

```{r results = "hide"}
sample = mcmc.spline(m, it = 3000, burning = 1000, thinning = 10)
```

As the mcmc function samples directly from the posterior, the result is a sample 
for the beta and gamma parameters as well as for the smoothing values. The
prediction with this sample is in result much smoother:

```{r}
summary(sample)
plot(sample, m)
```
To investigate properties of the sample we can now look at the random walk for different 
parameters.

```{r}
par(mfrow = c(2, 1))
par(mar = c(3, 5, 2, 4))
index = seq(1, length(sample$beta[,1]))
plot(index, sample$beta[,1], type = "l", xlab = "")
plot(index, sample$gamma[,1], type = "l", xlab = "")
plot(index, sample$epsilon, type = "l", xlab = "")
plot(index, sample$tau, type = "l")
```

While there seems to be no getting stuck, it seems that consecutive iterations
are correlated. We now could choose a higher thinning value or
alternatively set a custom stepsize for the sampling of the gamma values 
for the MCMC function.

